---
title: "Problem Set 36"
author: "Eric Zhou"
date: "2023-05-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **Problem 1**
#### **Part a**: Yes
#### **Part b**: 
```{r}
faithful.lm <- lm(faithful$waiting~faithful$eruptions)
res <- faithful$waiting - faithful.lm$fitted
res[1:10]
faithful.lm$residuals[1:10]
res-faithful.lm$residuals
```

#### **Part c**: Seems to be two clumps around x-values which is determined by data concentration.
```{r}
plot(faithful$eruptions,res)
```

#### **Part d**: Looks pretty good.
```{r}
hist(res)
```

<br>

## **Problem 2**
#### **Part a**: Actual data is the fitted data plus the residuals which is the error/offset
#### **Part b**: The larger the sum of squared residuals, the poorer the fit.
#### **Part c**: Shows how far off each prediction by the linear model is.
<br>

## **Problem 3**
#### **Part a**: 
```{r}
10.7296/0.3148
summary(faithful.lm)
```

#### **Part b**: Reject
#### **Part c**: 
```{r}
10.7296 + qt(0.975, 270) * 0.3148
10.7296 - qt(0.975, 270) * 0.3148
```

#### **Part d**: 
```{r}
1-var(faithful.lm$residuals)/var(faithful$waiting)
cor(faithful$waiting,faithful$eruptions)^2
```

