---
title: "Problem Set 30"
author: "Eric Zhou"
date: "2023-03-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **Problem 1**

#### **Part a**: $H_0: p \geq 0.51$, $H_A: p<0.51$

#### **Part b**:

```{r}
se <- sqrt(0.51*0.49/200)
se

x <- seq(from=0.51-3*se, to=0.51+3*se, by=0.001)
y <- dnorm(x, 0.51, se)
plot(x, y, type="l", xlab="bar X", ylab="probability")
abline(v=qnorm(0.05, 0.51, se), col="red")
```

#### **Part c**: 0.45 would be slightly smaller than p-value.

```{r}
qnorm(0.05, 0.51, se)
pnorm(0.45, 0.51, se)
```

<br>

## **Problem 2**

#### **Part a**: $H_0: \mu \leq 68$, $H_A: \mu>68$

#### **Part b**:

```{r}
se <- 21/sqrt(100)
se
x <- seq(from=68-3*se, to=68+3*se, by=0.01)
y <- dnorm(x, 68, se)
plot(x, y, type="l", xlab="bar X", ylab="probability")
abline(v=qnorm(0.9, 68, se), col="red")
```

#### **Part c**: Fail to reject null hypothesis

```{r}
qnorm(0.9, 68, se)
pnorm(69.2, 68, se)
```

<br>

## **Problem 3**

#### **Part a**: Null would be true mean of drug value is greater than or equal to some value.

#### **Part b**: Null would be proportion giving to charity is less than or equal to some value.

#### **Part c**: Null would be stretching does not affect flexibility, not one-sided.

#### **Part d**: Null would be true mean of sleep is greater than or equal to some value.

<br>

## **Problem 4: There was an extremely large effect in the opposite direction.**
