---
title: "Problem Set 33"
author: "Eric Zhou"
date: "2023-03-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **Problem 1**

#### **Part a**: $H_0:\mu_1=\mu_2=\mu_3=\mu_4, H_A:$ At least one mean is different

#### **Part b**: 4 airlines, 100 samples?

#### **Part c**: F-distribution

#### **Part d**: 4.711

#### **Part e**: 0.004, at least one mean is different.

<br>

## **Problem 2**

#### **Part a**: Character for airline, number for time.

```{r}
data <- read.csv("C:/Users/ezhou/OneDrive - Eastside Preparatory School/Documents/Eric/2022 - 2023/3 Spring/D Stats/airdelays.csv")
summary(data)
```

#### **Part b**: 5 airlines, 129 data entries.

```{r}
unique(data$airline)
```

#### **Part c**:

```{r}
boxplot(time~airline, data=data)
```

#### **Part d**: Aligns with boxplot very well.

```{r}
for (a in unique(data$airline)){
  print(a)
  print(mean(data$time[data$airline == a]))
}
```

<br>

## **Problem 3**

#### **Part a**:

```{r}
summary(aov(time~airline, data=data))
```

#### **Part b**: Reject null hypothesis, at least one is different.

#### **Part c**: H and Am but they are all kind of all over the place.
