---
title: "Problem Set 32"
author: "Eric Zhou"
date: "2023-03-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Difference of Means

<br>

## **Problem 1**

#### **Part a**: $H_0: \mu_1=\mu_2, H_A: \mu_1\neq\mu_2$. Fail to reject null hypothesis since p-score is higher than 0.05.

```{r}
pe <- 7.2 - 8.8
pe
se <- sqrt(1.35^2/6 + 2.46^2/5)
se
ts <- pe/se
ts
2 * pt(ts, 4)
```

#### **Part b**: Look at 95% confidence interval.

```{r}
pe + qt(0.025, 4) * se
pe + qt(0.975, 4) * se
```

<br>

## **Problem 2**

```{r}
Walnut_Park <- c(24.5, 12.6, 11, 13, 11.1, 19.1, 20.6, 16.8, 11.5, 16.2, 24.9, 19.9)
Sunnyside <- c(10.5, 14.6, 13.8, 5.3, 7.6, 7.5, 10.3, 10.2, 6.4, 9.2, 22.1, 14.3)
length(Walnut_Park)
length(Sunnyside)
mean(Walnut_Park)
mean(Sunnyside)
```

#### **Part a**: Reject null hypothesis with p-value of only around 0.014.

```{r}
pe <- mean(Walnut_Park) - mean(Sunnyside)
pe
se <- sqrt(sd(Walnut_Park)^2/12 + sd(Sunnyside)^2/12)
se
ts <- pe/se
ts
2 * pt(ts, 11, lower.tail=F)
t.test(Walnut_Park, Sunnyside)
```

#### **Part b**:

```{r}
pe + qt(0.05, 11) * se
pe + qt(0.95, 11) * se
```

<br>

# Paired Data

<br>

## **Problem 3:** $H_0: \mu_{diff}=0, H_A: \mu_{diff}\neq0$. Fail to reject null hypothesis since p-score is higher than 0.05.

```{r}
ts <- 2.18/11.45
ts
2 * pt(ts, 27, lower.tail=F)
```

<br>

## **Problem 4: Reject cuz 0.05 \< 0.1 alpha value meaning there is a significant change.**

```{r}
before <- c(5, 12, 8, 5, 7, 5, 7, 4, 6, 5, 8, 11)
after <- c(6, 7, 6, 3, 6, 4, 3, 8, 5, 4, 4, 9)
length(before)
```

```{r}
diffs <- before - after
pe <- mean(diffs)
pe
se <- sd(diffs)/sqrt(12)
se
ts <- pe/se
ts
2 * pt(ts, 11, lower.tail = F)
t.test(before, after, pair=T)
```
