---
title: "Problem Set 13"
author: "Eric Zhou"
date: "2022-10-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **Problem 1**
```{r}
setwd("C:/Users/ezhou/OneDrive - Eastside Preparatory School/Documents/Eric/2022 - 2023/1 Fall/D Stats")
sounders <- read.csv("sounders.csv")
```

#### **Part a**: Data is 34 x 8.  The 8 columns are 'Date', 'Opponent', 'Goals_For', 'Goals_Against', 'WDL', 'Home_Away', 'Attendance', and 'Possession'
```{r}
dim(sounders)
head(sounders)
```

#### **Part b**: Put up 52 goals total
```{r}
sum(sounders$Goals_For)
```

#### **Part c**: Won 16 games
```{r}
length(which(sounders$Goals_For > sounders$Goals_Against))
length(which(sounders$WDL == 'W'))
```

#### **Part d**: Mean is 1.529412, median is 1
```{r}
mean(sounders$Goals_For)
median(sounders$Goals_For)
```

#### **Part e**: Mean is 29316.35, median is 31967.5. The difference between mean and medium can show the variance and so does the histogram. I would say median is a better estimate here because there's a lot of games around there and mean combines the outliers on both ends and ends up in the middle with no data values
```{r}
mean(sounders$Attendance)
median(sounders$Attendance)
hist(sounders$Attendance)
```

#### **Part f**: 
```{r}
quantile(sounders$Possession)
```

<br>

## **Problem 2**
#### **Part a**:Concentrated at the 0 - 1 and then decreases as the score increases
```{r}
hist(sounders$Goals_For, xlab='Goals Scored', main='Sounders Goals Scored', breaks=3)
```

#### **Part b**: There's a dip in the middle and then a dip at the end. Overall seems pretty random and jumps around a lot. Used lines to track pattern
```{r}
plot(x=sounders$Goals_For, type='l')
```

#### **Part c**: Spaced out, not a normal distribution. There's 2 higher ones and everything else is really infrequent
```{r}
hist(sounders$Attendance, xlab='Attendance', main='Sounders Games Attendance', breaks=10)
```

<br>

## **Problem 3**
#### **Part a**: 
```{r}
sounders[sounders$Attendance > 38000, ]
```

#### **Part b**: 
```{r}
sounders[sounders$WDL == 'D', ]
```

#### **Part c**: 
```{r}
sounders[sounders$Goals_For == 3 | sounders$Goals_For == 4, c('Date', 'Opponent')]
```

#### **Part d**: ~20.59% they were completely shut out
```{r}
length(sounders[sounders$Goals_For == 0,1])/length(sounders[,1])
```

<br>

## **Problem 4**
#### **Part a**: 
```{r}
goal_diff = sounders$Goals_For - sounders$Goals_Against
goal_diff
```

#### **Part b**: They would get 56 points total
```{r}
length(which(sounders$WDL == 'W')) * 3 + length(which(sounders$WDL == 'D'))
```

