---
title: "Problem Set 12"
author: "Eric Zhou"
date: "2022-10-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **Problem 1**
```{r}
temps <- c(71, 71, 69, 71, 69, 72, 71, 70, 70, 71, 82, 76, 74, 73, 74, 82, 82, 77, 80, 83, 83, 81, 72, 72, 71, 72, 72, 71, 72, 72, 73)
```

#### **Part a**: 5.5 degrees
```{r}
IQR(temps)
```

#### **Part b**: 71 to 76.5 degrees
```{r}
quantile(temps, c(0.25, 0.75))
```

#### **Part c**: IQR is the difference between the top and lower bound of the middle 50%
#### **Part d**: Looks like there are no outliers
```{r}
boxplot(temps, ylab="Temperature per Year")
```

<br>

## **Problem 2**
```{r}
section1 <- c(93, 89, 99, 95, 78, 81, 83, 94, 75, 98, 87, 94, 96, 94, 89)
sort(section1)
```

#### **Part a**: 13.33%
```{r}
sum(section1 < 80)/length(section1)
```

#### **Part b**: 13.33%
```{r}
sum(section1 >= 98)/length(section1)
```

#### **Part c**: 20%
```{r}
sum(section1 == 94)/length(section1)
```

#### **Part d**: Student #8, #12, and #14
```{r}
studs = seq(1, length(section1))
studs[section1 == 94]
```

#### **Part e**: 79.2
```{r}
quantile(section1, 0.1)
```

#### **Part f**: 93
```{r}
quantile(section1, 0.5)
```

<br>

## **Problem 3**
#### **Part a**: 
```{r}
hist(temps)
abline(v=quantile(temps, 0.05), col='green')
abline(v=quantile(temps, 0.95), col='green')
```

#### **Part b**: 
```{r}
hist(temps, xlim = c(65, 85))
abline(v=quantile(temps, 0.05), col='green')
abline(v=quantile(temps, 0.95), col='green')
abline(v=mean(temps)-1.68*sd(temps), col='blue')
abline(v=mean(temps)+1.68*sd(temps), col='blue')

```

#### **Part c**: Part a does since it actually uses what the 25% and 75% values of the scores are

<br>

## **Problem 4**
```{r}
enrollment <- c(6414, 1550, 2109, 9350, 21828, 4300, 5944, 5722, 2825, 2044, 5481, 5200, 5853, 2750, 10012, 6357, 27000, 9414, 7681, 3200, 17500, 9200, 7380, 18314, 6557, 13713, 17768, 7493, 2771, 2861, 1263, 7285, 28165, 5080, 11622)
sort(enrollment)
```

#### **Part a**: There are outliers. They are 21828, 27000, and 28165
```{r}
boxplot(enrollment, ylab='Students Enrolled', main='Community College Enrollment')
```

#### **Part b**: I would want to accomodate the middle 50%, so 3750 to 9713 students. I think that makes sense since it's the middle half so not too little, not too much.
```{r}
quantile(enrollment, c(0.25, 0.75))
```

