---
title: "Problem Set 11"
author: "Eric Zhou"
date: "2022-10-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## **Problem 1**
```{r}
section1 <- c(92, 88, 100, 96, 80, 82, 85, 93, 81, 95, 84, 92, 93, 90, 92)
section2 <- c(92, 90, 99, 95, 78, 81, 83, 94, 75, 98, 94, 96, 94, 89)
```

#### **Part a**: 
```{r}
var(section1)  # section 1 variance
sd(section1)  # section 1 standard dev
var(section2)  # section 2 variance
sd(section2)  # section 2 standard dev
```

#### **Part b**: Unit for variance is points/percentage squared, unit for standard deviation is points/percentage
#### **Part c**: Section 2 has a larger standard deviation (7.63 compared to 5.96)
#### **Part d**: Section 1 is 77.61 to 101.46, section 2 is 74.59 to 105.13
```{r}
mean(section1) - 2 * sd(section1)
mean(section1) + 2 * sd(section1)
mean(section2) - 2 * sd(section2)
mean(section2) + 2 * sd(section2)
```

#### **Part e**: Yes because variance is just standard deviation squared
<br>

## **Problem 2**
#### **Part a**: 
```{r}
enrollment <- c(6414, 1550, 2109, 9350, 21828, 4300, 5944, 5722, 2825, 2044, 5481, 5200, 5853, 2750, 10012, 6357, 27000, 9414, 7681, 3200, 17500, 9200, 7380, 18314, 6557, 13713, 17768, 7493, 2771, 2861, 1263, 7285, 28165, 5080, 11622)
```

#### **Part b**: 
```{r}
hist(enrollment, breaks=10)
```

#### **Part c**: Still not sure which one is better. Mean is skewed by outliers on either end but median is not good for representing the data as a whole but does get the middle. For this, I think I would say median is better but honestly no clue.
```{r}
mean(enrollment)
median(enrollment)
```

#### **Part d**: Unit is number of people
```{r}
sd(enrollment)
```

<br>

## **Problem 3**
#### **Part a**: 
```{r}
up_bound = mean(enrollment) + sd(enrollment)
low_bound = mean(enrollment) - sd(enrollment)
up_bound
low_bound
```

#### **Part b**: 
```{r}
hist(enrollment, breaks=10)
abline(v=up_bound, col='orange')
abline(v=low_bound, col='orange')
```

#### **Part c**: 27 schools are within the range
```{r}
range_enroll <- enrollment[enrollment > low_bound]
range_enroll <- range_enroll[range_enroll < up_bound]
range_enroll
length(range_enroll)
```


#### **Part d**: Max: 28165, Min: 1263, Range: 26902
```{r}
max(enrollment)
min(enrollment)
max(enrollment) - min(enrollment)
```

#### **Part e**: Probably somewhere betwen 1500 and 20000
<br>

## **Problem 4**:
```{r}
sd(section1)
diffs <- (section1 - mean(section1))^2
sqrt(sum(diffs)/(length(diffs) - 1))
```

