---
title: "Assignment 7"
author: "Eric Zhou"
date: "2022-12-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# **Apprentice**
## Problem 1
#### **Part a**: Random variable would be the result of the 2 die.
```{r}
sum(sample(1:4, 1), sample(1:4, 1))
```

#### **Part b**: Total chance of winning $4 is 7/16, getting money back is 1/4, and losing $5 is 5/16.
Sum | Probability | Outcome
:-:|:-:|:-:
2 | 1/16 | 0
3 | 2/16 | -5
4 | 3/16 | -5
5 | 4/16 | 4
6 | 3/16 | 4
7 | 2/16 | 0
8 | 1/16 | 0

#### **Part c**: E[X] is 0.1875.
```{r}
out <- c(0, -5, -5, 4, 4, 0, 0)
probs <- c(1/16, 2/16, 3/16, 4/16, 3/16, 2/16, 1/16)
ex <- sum(out * probs)
ex
```

#### **Part d**: Variance and standard deviation for one round of this game is approximately 14.78 and 3.84 respectively.
```{r}
varx <- sum(out^2*probs) - (ex)^2
varx
sqrt(varx)
```

#### **Part e**: 120 rounds would have expected winnings of $22.5 with a standard deviation of $42.11
```{r}
mulex <- ex * 120
mulsd <- sqrt(120 * varx)
mulex
mulsd
```

#### **Part f**: 95% likely range for 120 rounds is from -$61.72 to 106.72
```{r}
mulex + 2 * mulsd
mulex - 2 * mulsd
```

#### **Part g**: I would since the game is skewed to the positive so the chance of winning money is higher than that of losing money.
<br>

# **Master**
## Problem 2
#### **Part a**: Around $3.57
```{r}
out2 <- c(0, -5, -5, 0, 0)
probs2 <- c(1/16, 2/16, 3/16, 2/16, 1/16)
-sum(out2 * probs2)/(7/16)
```

#### **Part b**: Z-score of 0 is -0.53, meaning that the chance of losing money is quite a bit lower than even and that 69.1% of the distribution is above 0. This is a lot lower than the z-score for a single round and that makes sense because the chance of earning money is higher so as you do more, you should theoretically earn more.
```{r}
(0-mulex)/mulsd
(0-ex)/sqrt(varx)
```

#### **Part c**: Risk prone is willing to take risks to gain money, risk neutral is where risk is not considered while gaining money, risk averse is to avoid risks. I think this I would classify myself as between risk neutral and risk averse but leaning towards averse since I chose to play the game only because the expected value is positive and the z-score of zero leans towards not losing money. However, there's obviously still a chance to lose money so I am not weighting that as heavily.
#### **Part d**: This simulation of 500 rounds averaged $0.166 of earnings per round and had a standard deviation of around $3.85, which is decently close to the $0.1875 and $3.84.
```{r}
earnings <- c()
for (j in 1:500){
  roll <- sum(sample(1:4, 1), sample(1:4, 1))
  earnings[j] <- out[roll - 1]
}
mean(earnings)
sd(earnings)
``
<br>

## Problem 3
#### **Part a**: Maximum earnings is $700, minimum is -$80
```{r}
60 * 15 - 200
8 * 15 - 200
```

#### **Part b**: Expected value for number of people is 35 and expected profit for the fundraiser is $325.
```{r}
expeep <- sum(60 * 0.45 + 20 * 0.3 + 8 * 0.25)
expeep
expeep * 15 - 200
```

#### **Part c**: I would say no because although the expected value is positive, the fundraiser is only going to occur once so you can't benefit from doing multiple rounds. Furthermore, there's a 25% chance that you do work and lose money which seems pretty sad. I would try to do a different fundraiser that guarantees no losses. 
<br>

# **Wizard**
## Problem 4
#### The simulation had average earnings of $22.62 with a standard deviation of around $41.28. This closely aligns with the values calculated earlier which was an expected value of $22.5 with a standard deviation of $42.11.
```{r}
earnings <- c()
for (j in 1:500) {
  rolls <- sample(1:4, 120, replace=TRUE) + sample(1:4, 120, replace=TRUE)
  earnings[j] <- sum(out[rolls - 1])
}
mean(earnings)
sd(earnings)
```

<br>

## Problem 5
#### **Part a**: Expected value for loaves sold is 3.05 with an expected earning of $14.49.
```{r}
loafprobs <- c(0.05, 0.15, 0.2, 0.2, 0.2, 0.1, 0.1)
exloaf <- sum(0:6 * loafprobs)
exloaf
exloaf * 4.75
```

#### **Part b**: Standard deviation for revenue is around $7.87.
```{r}
sqrt(sum((0:6)^2 * loafprobs) - (exloaf)^2) * 4.75
```

#### **Part c**: 
```{r}
sum(c(0, 1, 2, 3, 4, 5, 5) * loafprobs) * 4.75 - 5 * 1.75
```

#### **Part d**: 4 loaves has the highest expected profit.
```{r}
for (j in 0:6) {
  bread <- c(0:6)
  for (v in seq(length(bread))) {
    bread[v] = min(v - 1, j)
  }
  print(sum(bread * loafprobs) * 4.75 - j * 1.75)
}
```

#### **Part e**: Size of the oven and if you could store the loaves for later since then it ight be more efficient just to bake more. Also, maybe bulk buying ingredients.
