---
title: "Assignment 11"
author: "Eric Zhou"
date: "2023-02-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# **Apprentice**

## Problem 1

#### **Part a**: $a=0, b=50$

#### **Part b**: Mean is 25, variance is 208.33, sd is 14.43.

```{r}
(0+50)/2
(50-0)^2/12
sqrt((50-0)^2/12)
```

#### **Part c**: 0.02 but that's a tiny slice of the total and not very helpful when looking at continuous uniform distributions. Generally you look at a range.

```{r}
dunif(24.2, 0, 50)
```

#### **Part d**: 0.62 chance of W being less than 31.

```{r}
punif(31, 0, 50)
```

#### **Part e**: 40% chance between 27 and 47.

```{r}
punif(47, 0, 50) - punif(27, 0, 50)
```

#### **Part f**: Mean was within 0.1 of the prediction and variance was within 5.

```{r}
vals <- runif(1000, 0, 50)
mean(vals)
var(vals)
```

<br>

## Problem 2

#### **Part a**: $\mu=16.8, \sigma=2.7$

#### **Part b**: Mean and standard deviation match really well which makes sense.

```{r}
vals <- rnorm(1000, 16.8, 2.7)
mean(vals)
sd(vals)
hist(vals, freq=F, main="Y~N(16.8, 2.7)", xlab="Value", ylab="Probability")
```

#### **Part c**:

```{r}
x <- seq(16.8 - 3 * 2.7, 16.8 + 3 * 2.7, by=0.01)
hist(vals, freq=F, main="Y~N(16.8, 2.7)", xlab="Value", ylab="Probability")
lines(x, dnorm(x, 16.8, 2.7), col="blue", type="l")
```

#### **Part d**: 74.75% is below 18.6

```{r}
pnorm(18.6, 16.8, 2.7)
hist(vals, freq=F, main="Y~N(16.8, 2.7)", xlab="Value", ylab="Probability")
lines(x, dnorm(x, 16.8, 2.7), col="blue", type="l")
abline(v=18.6, col="red")
```

#### **Part e**: Only 5.99% above 21.

```{r}
pnorm(21, 16.8, 2.7, lower.tail=F)
hist(vals, freq=F, main="Y~N(16.8, 2.7)", xlab="Value", ylab="Probability")
lines(x, dnorm(x, 16.8, 2.7), col="blue", type="l")
abline(v=21, col="red")
```

#### **Part f**: 21.04% of data is between 18 and 20.

```{r}
pnorm(20, 16.8, 2.7) - pnorm(18, 16.8, 2.7)
```

#### **Part g**: 19.60 is above 85% of the distribution

```{r}
qnorm(0.85, 16.8, 2.7)
```

#### **Part h**: 12.91 to 20.69 is the middle 85%

```{r}
qnorm(0.075, 16.8, 2.7)
qnorm(0.85 + 0.075, 16.8, 2.7)
```

<br>

# **Master**

## Problem 3

```{r}
a <- c(-0.12,-0.39,-0.73,-0.67,-0.44,0.15,-0.69,-1.3,-0.66,-0.58,-0.44,0.3,-0.36,-0.29,-0.52,-0.21,0.02,-0.34,0.21,-0.58,0.08,0.89,-0.31,-0.42,0.36,-0.34,-0.46,0,0.55,0.24,0.34,0.52,0.31,0.19,0.48,-0.1,0.13,0.03,-0.19,-0.06,-0.63,-0.49,0.42,0.72,0.34,-0.27,-0.31,0.13,0.31,-0.16,0.09,-0.12,0.03,0.19,-0.36,-0.42,-0.41,-0.2,-0.16,-0.07,-0.27,-0.3,-0.92,0.22,-0.28,-0.3,-0.05,0.56,-0.21,-0.28,0.31,1.13,-0.57,0.1,0.11,-0.17,0.46,1.13,0.48,-0.04,0.31,0.55,0.15,0.22,0.4,0.43,-0.22,0.45,1.55,1.09,0.77,1.19,0.69,0.88,0.52,1.24,1.5,0.83,0.29,0.4,1.44,0.92,1.54,0.53,0.59,1.38,1.93,1.5,0.88,0.92)
```

#### **Part a**: Mean is 0.175, standard deviation is 0.612.

```{r}
temp_mean <- mean(a)
temp_sd <- sd(a)
temp_mean
temp_sd
```

#### **Part b**:

```{r}
x <- seq(temp_mean - 3 * temp_sd, temp_mean + 3 * temp_sd, by=0.01)
hist(a, freq=F, main="North America Temperature Difference Distribution", xlab="Temperature Difference", ylab="Probability")
lines(x, dnorm(x, temp_mean, temp_sd), col="blue", type="l")
```

#### **Part c**: It looks like a pretty decent fit for the data.

#### **Part d**: 61% of temperature anomalies are increases in temperature from average.

```{r}
pnorm(0, temp_mean, temp_sd, lower.tail=F)
```

#### **Part e**: Original data has 56.36% above 0.

```{r}
length(a[a>0])/length(a)
```

#### **Part f**: Very close, with 11.60% as prediction and 11.82% as actual data.

```{r}
pnorm(-1, temp_mean, temp_sd) + pnorm(1, temp_mean, temp_sd, lower.tail=F)
(length(a[-1>a]) + length(a[a>1]))/length(a)
```

<br>

## Problem 4

```{r}
x <- c(4, 5, 6, 7, 8, 9)
p <- c(0.01, 0.09, 0.69, 0.15, 0.04, 0.02)
```

#### **Part a**:

```{r}
barplot(p ~ x, ylab="Probability", xlab="Value", main="Trimester Classes Distribution")
```

#### **Part b**: Expected value is 6.18 and variation 0.5876.

```{r}
ev <- sum(x*p)
ev
ev2 <- sum(x^2*p)
varx <- ev2-ev^2
varx

```

#### **Part c**:

```{r}
vals <- c()
for (j in 1:4000){
  vals[j] <- sum(sample(x, 12, prob=p, replace=T))
}
hist(vals, freq=F, ylab="Probability", xlab="Classes", main="Total Classes Distribution")
```

#### **Part d**: Average is 74.13 and standard deviation is 2.65.

```{r}
mean(vals)
sd(vals)
```

#### **Part e**: Aligns really well with what we found in part d.

```{r}
mean12 <- 12 * ev
sd12 <- sqrt(12) * sqrt(varx)
x <- seq(mean12 - 3 * sd12, mean12 + 3 * sd12, by=0.01)
hist(vals, freq=F, ylab="Probability", xlab="Classes", main="Total Classes Distribution")
lines(x, dnorm(x, mean12, sd12), col="blue", type="l")
```

#### **Part f**: It is definitely working. The new distribution fits the normal curve a lot better than what we started with.
