---
title: "Problem Set 26"
author: "Eric Zhou"
date: "2023-02-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **Problem 1**

#### **Part a**: $a=3, b=8$

#### **Part b**:

```{r}
(8+3)/2
(8-3)^2/12
sqrt((8-3)^2/12)
```

#### **Part c**: A very small amount because you care more about ranges.

#### **Part d**:

```{r}
punif(5.42, 3, 8)
```

#### **Part e**:

```{r}
punif(7.21, 3, 8) - punif(3.48, 3, 8)
```

#### **Part f**: Matches really well

```{r}
thing <- runif(2000, 3, 8)
var(thing)
mean(thing)
```

<br>

## **Problem 2**

#### **Part a**: Areas of ranges times densities

#### **Part b**: They are both uniform but one has values with decimals and includes all numbers while the other includes certain sets of numbers.

#### **Part c**: Linearly increasing.

<br>

## **Problem 3**

#### **Part a**:

```{r}
hist(runif(1000, 0, 1))
```

#### **Part b**:

```{r}
thing <- runif(1000, 0, 1)
thing2 <- thing * 5 + 3
mean(thing2)
var(thing2)
```

#### **Part c**: \* 50

#### **Part d**: Not sure what was wanted here

```{r}
qbinom(runif(1, 0, 1), 18, 1)
```

<br>

## **Problem 4**

```{r}
(5-0)/(10-0)
qunif(5/10, 0, 10)
```
