---
title: "Problem Set 22"
author: "Eric Zhou"
date: "2023-01-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **Problem 1**
#### **Part a**: Expect 21.25, sd of around 1.79
```{r}
0.85 * 25
sqrt(25*0.85*0.15)
```

#### **Part b**: n = 25, k = 18+, p = 0.85
```{r}
1 - pbinom(18 - 1, 25, 0.85)
```

#### **Part c**: 
```{r}
pbinom(18, 25, 0.85)
```

#### **Part d**: No, because x = 18 is repeated in both.
#### **Part e**: Between 17.75 and 24.75 which aligns well with the 18 earlier.
```{r}
0.85 * 25 + 1.96 * sqrt(25*0.85*0.15)
0.85 * 25 - 1.96 * sqrt(25*0.85*0.15)
```

<br>

## **Problem 2**
#### **Part a**: 
```{r}
dbinom(10, 10, 94/117)
1 - pbinom(9, 10, 94/117)
```

#### **Part b**: 
```{r}
pbinom(3, 5, 94/117)
```

#### **Part c**: 
```{r}
1 - pbinom(4 - 1, 5, 94/117)
```

#### **Part d**: 
```{r}
1 - pbinom(8, 10, 94/117)
```

#### **Part e**: I assume so but in real life probably not since skill level would affect each player and team's success rate.
#### **Part f**: 95% likely range.
```{r}
10 * 94 / 117 + 1.96 * sqrt(10 * 94 / 117 * 23 / 117)
10 * 94 / 117 - 1.96 * sqrt(10 * 94 / 117 * 23 / 117)
```

#### **Part g**: 
```{r}
probs <- dbinom(0:5, 5, 94/117)
probs
sum(probs ^ 2)
```

<br>

## **Problem 3**
#### **Part a**: 
```{r}
probs <- pbinom(0:25, 25, 0.85)
probs
```

#### **Part b**: 
```{r}
plot(probs, type="s")
```

#### **Part c**: Adds cumulative probabilities at each step.
#### **Part d**: Both are at 23, makes sense since it adds enough to include both thresholds

