---
title: "Problem Set 25"
author: "Eric Zhou"
date: "2023-01-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **Problem 1**
#### **Part a**: Estimate for true difference in proportions so difference is 28% of people.
```{r}
6/20 - 11/19
```

#### **Part b**: H_0: p1 = p2, H_A : p1 != p2, fail to reject H_0
```{r}
prop.test(c(14, 8), c(20, 19), correct=F)
```

#### **Part c**: 0.158859
```{r}
p <- 22/39
sqrt(p * (1-p) * (1/20 + 1/19))
```

#### **Part d**: Although we fail to reject and they should be identically distributed, it does not look to be that way. Independence is the same.
```{r}
barplot(matrix(c(6, 14, 11, 8),nr=2), beside=T, col=c("aquamarine3","coral"), 
        names.arg=c("Latin", "No Latin"), ylim=c(0, 15))
legend("topleft", c("<= 100","> 100"), pch=15,  col=c("aquamarine3","coral"), 
       bty="n")
```

#### **Part e**: -0.02 to 0.58. Does include 0 meaning it is a possible difference between the 2 so we would fail to reject.

<br>

## **Problem 2**
#### **Part a**: No, p-value is around 0.1.
#### **Part b**: Contains 0 so we would fail to reject as it is a possibility in the 95% range.