---
title: "Problem Set 24"
author: "Eric Zhou"
date: "2023-01-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **Problem 1**

#### **Part a**: $n=200, Y=120, \hat{p}=0.6$

#### **Part b**:

```{r}
se <- sqrt(0.6*0.4/200)
se
```

#### **Part c**:

```{r}
0.6 - 1.96 * se
0.6 + 1.96 * se
```

#### **Part d**:

```{r}
1.96 * se
```

#### **Part e**:

```{r}
n2 <- 1/sqrt(0.025/sqrt(0.6*0.4))
n2
sqrt(0.6*0.4/n2)
```

<br>

## **Problem 2**

#### **Part a**: 95% of confidence intervals created this way will include true mean

#### **Part b**: Standard deviation is distance of values from the mean, standard error is difference between sample p_hat and population actual p_hat

#### **Part c**: Standard error decreases as population increases

<br>

## **Problem 3**

#### **Part a**:

```{r}
sqrt(0.43*(1-0.43))/sqrt(600)
```

#### **Part b**:

```{r}
sqrt(0.43*(1-0.43))/sqrt(1200)
```

#### **Part c**: Divided by square root 2s

```{r}
sqrt(2)
```

#### **Part d**: 2400 cuz sqrt(4) = 2

```{r}
sqrt(0.43*(1-0.43))/sqrt(2400)
```

<br>

## **Problem 4**

#### **Part a**:

```{r}
se <- sqrt(0.51*0.49/600)
0.51 - 1.64 * se
0.51 + 1.64 * se
```

#### **Part b**:

```{r}
0.51 - 2.58 * se
0.51 + 2.58 * se
```

#### **Part c**: Part b because it encompasses a larger percentage.
