---
title: "Problem Set 21"
author: "Eric Zhou"
date: "2023-01-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **Problem 1**
#### **Part a**: n is 32, p is 36%, k depends on how many you are looking for to say that social media is their primary news source.
#### **Part b**: Around a 10% chance.
```{r}
choose(32, 9)*(0.36)^9*(0.64)^23
dbinom(9, 32, 0.36)
```

#### **Part c**: Expect around 11.52 people so round up to 12.
```{r}
0.36*32
```

#### **Part d**: Around 0.00006% which is very small. This makes sense since it's still possible for it to happen (64% raised to 32 power), it's just not a large chance.
```{r}
dbinom(0, 32, 0.36)
0.64^32
```

#### **Part e**: 
```{r}
barplot(table(rbinom(10000, 32, 0.36)))
```

<br>

## **Problem 2**
#### **Part a**: n is 20, k is 11, p is 0.4. Around 7.10%
```{r}
dbinom(11, 20, 0.4)
```

#### **Part b**: Expect 8 with standard deviation of 2.19.
```{r}
20 * 0.4
sqrt(20 * 0.4 * 0.6)
```

#### **Part c**: ~87.25%
```{r}
tot <- 0
for (i in 0:10){
  tot <- tot + dbinom(i, 20, 0.4)
}
tot
```

#### **Part d**: 12.75%
```{r}
tot <- 0
for (i in 11:20){
  tot <- tot + dbinom(i, 20, 0.4)
}
tot
```

<br>

## **Problem 3**
#### **Part a**: 1 and 6
#### **Part b**: 
```{r}
(1 + 6) / 2
sum(c(1:6)/6)
```

#### **Part c**: 
```{r}
((6-1+1)^2-1)/12
ex <- sum(c(1:6)/6)
ex2 <- sum(c(1:6)^2/6)
sum(ex2 - ex^2)
```

<br>

## **Problem 4**
#### **Part a**: Normal distribution concentrated at the middle.
#### **Part b**: Choosing 0 from n, all false, does not affect distribution. It's just the very first bar.
#### **Part c**: Choosing n from n, all true, does not affect distribution. It's just the very last bar.
#### **Part d**: You are just summing up every possibility which adds up to 1.