---
title: "Problem Set 15"
author: "Eric Zhou"
date: "2022-11-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **Problem 1**
#### **Part a**: Expected value is 4.5
```{r}
sum(seq(1, 8)/8)
```

#### **Part b**: 4.5 is not a value that can occur because it is not one of the sides of the die, which are whole numbers from 1 to 8 inclusive
#### **Part c**: A list of the results of 1000 different rolls of a fair 8-sided die
```{r}
krolls <- sample(seq(1, 8), replace = TRUE, 1000)
krolls
```

#### **Part d**: 
```{r}
barplot(table(krolls))
abline(h=1000/8)
```

#### **Part e**: They are pretty close. tEh statistical average is 4.5 while the 1000 trials' average is 4.566
```{r}
sum(krolls)/1000
```

<br>

## **Problem 2**
#### **Part a**: 
Outcome | Winnings ($) | Probability
---|:-:|:-:
number card | 0 | 0.69
face card | 3 | 0.23
Ace (no Spades) | 5 | 0.06
Ace of Spades | 25 | 0.02

#### **Part b**: Yes because all the cards together constitutes 52/52, which is 1.
#### **Part c**: Expected value of winning is $1.49 per round.
```{r}
wins <- c(0, 3, 5, 25)
chances <- c(0.69, 0.23, 0.06, 0.02)
sum(wins * chances)
```

#### **Part d**: Yes because for a price of $2, you earn an average of $1.49 back per round, that's almost a 75% profit.

<br>

## **Problem 3**
#### **Part a**: A is $120 net, B is -$50 net, and C is perfectly even
```{r}
(400 * 0.8 + 0 * 0.2) - 200
(600 * 0.5 + 100 * 0.5) - 400
(1500 * 0.25 + 500 * 0.75) - 750
```

#### **Part b**: I would choose A since it's the only expected value that profits
#### **Part c**: What the companies do and whether that aligns with my values. Maybe my personal beliefs regarding the topic rather than relying purely on statistics.
<br>

## **Problem 4**
#### **Part a**: 
```{r}
sample(seq(1, 13), 1)
```

#### **Part b**: 
```{r}
card <- sample(seq(1, 13), 1)
if (card == 1) {
  print(paste("Ace:", card))
} else if (card >= 11) {
  print(paste("Face card:", card))
} else {
  print(paste("Number card:", card))
}
```

#### **Part c**: 
```{r}
card <- sample(seq(1, 13), 1)
if (card == 1) {
  if (sample(seq(1, 4), 1) == 1) {
    print(paste("Ace of Spades:", card, "| Net Earnings:", 23))
  } else {
    print(paste("Ace (not Spades):", card, "| Net Earnings:", 3))
  }
} else if (card >= 11) {
  print(paste("Face card:", card, "| Net Earnings:", 1))
} else {
  print(paste("Number card:", card, "| Net Earnings:", -2))
}
```

